### 2021-03-31 ch05_paper_ecs/paper_ecs.tex
Further optimizations of abstract, introduction and conclusion

One of the most important sources of information for this purpose is
undoubtedly observational data. As presented in
\cref{subsec:02:emergent_constraints}, appropriate observations of the Earth
system can be utilized within the so-called \emph{emergent constraints}
framework to assess the accuracy of climate model projections and reduce
associated uncertainties. For \ac{ECS}, a large variety of different emergent
constraints has been developed over the years, which we now evaluate on the
\acs{CMIP}6 ensemble in order to assess whether they still hold for these new
climate models, \ie{} whether they have skill to reduce uncertainties in
\ac{ECS} in the \acs{CMIP}6 ensemble. In total we assess eleven emergent
constraints on \ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out
of which the most are related to cloud feedbacks since these constitute the
most important source of uncertainty for \ac{ECS} \commentcite{Boucher2013,
  Flato2013}{see also \cref{subsec:02:climate_feedbacks}}. Since all of the
evaluated emergent constraints have been derived on the \acs{CMIP}3 or
\acs{CMIP}5 ensemble, the \acs{CMIP}6 models offer an exciting possibility to
perform out-of-sample testing using the previously unseen data.

@@

In this \namecref{ch:05:paper_ecs}, eleven different emergent constraints on
\ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,
are assessed by applying them to results from \acp{ESM} contributing to
\acs{CMIP}5 and \acs{CMIP}6. Of particular interest are the results from
\acs{CMIP}6, since all analyzed emergent constraints have been published prior
to the availability of \acs{CMIP}6 data. In summary, the best estimate of
\ac{ECS} averaged over all emergent constraints increases by $12 \unit{\%}$
when moving from relationships trained on \acs{CMIP}5 to those trained on
\acs{CMIP}6. Some increase is predicted by almost every constraint we analyze
(with TII as the only exception) and can be at least partly explained by the
increased multi-model mean \ac{ECS} of \acs{CMIP}6, which has not been
accompanied by systematic changes in the constraint variables that could
explain this increase, leading to regression fits with higher intercept values
at observed constraint values. This is also illustrated by the \acs{CMIP}5 and
\acs{CMIP}6 \acp{MMM} in the left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
(colored circles), in which the connecting line between the \acs{CMIP}5 and
\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
for all emergent constraints.

>>>>

One of the most important sources of information for this purpose is
undoubtedly observational data. As presented in
\cref{subsec:02:emergent_constraints}, appropriate observations of the Earth
system can be utilized within the so-called \emph{emergent constraints}
framework to assess the accuracy of climate model projections and reduce
associated uncertainties. For \ac{ECS}, a large variety of different emergent
constraints has been developed over the years, which we now evaluate on the
\acs{CMIP}6 ensemble in order to assess whether they still hold for these new
climate models, \ie{} whether they have skill to reduce uncertainties in
\ac{ECS} in the \acs{CMIP}6 ensemble. In total we assess eleven emergent
constraints on \ac{ECS} (see \cref{tab:05:overview_emergent_constraints}), out
of which the most are related to cloud feedbacks since these constitute the
most important source of uncertainty for \ac{ECS} \commentcite{Boucher2013,
  Flato2013}{see also \cref{subsec:02:climate_feedbacks}}. Since all of the
evaluated emergent constraints have been derived from the \acs{CMIP}3 or
\acs{CMIP}5 ensemble, the \acs{CMIP}6 models offer an exciting possibility to
perform out-of-sample testing using the previously unseen data.

@@

In this \namecref{ch:05:paper_ecs}, eleven different emergent constraints on
\ac{ECS}, which are mostly directly or indirectly related to cloud feedbacks,
are assessed by applying them to results from \acp{ESM} contributing to
\acs{CMIP}5 and \acs{CMIP}6. Of particular interest are the results from
\acs{CMIP}6, since all analyzed emergent constraints have been published prior
to the availability of \acs{CMIP}6 data. In summary, the best estimate of
\ac{ECS} averaged over all emergent constraints increases by $12 \unit{\%}$
when moving from relationships trained on \acs{CMIP}5 to those trained on
\acs{CMIP}6. Some increase is predicted by almost every constraint we analyze
(with TII as the only exception) and can at least be partly explained by the
increased multi-model mean \ac{ECS} of \acs{CMIP}6, which has not been
accompanied by systematic changes in the constraint variables that could
explain this increase, leading to regression fits with higher intercept values
at observed constraint values. This is also illustrated by the \acs{CMIP}5 and
\acs{CMIP}6 \acp{MMM} in the left columns of
\cref{fig:05:bri_cox_lip,fig:05:shd_shl_shs,fig:05:su_tih_tii,fig:05:vol_zha}
(colored circles), in which the connecting line between the \acs{CMIP}5 and
\acs{CMIP}6 \ac{MMM} is not parallel to the \acs{CMIP}5 emergent relationships
for all emergent constraints.

### 2021-03-31 ch07_conclusion/conclusion.tex
Further optimizations of abstract, introduction and conclusion

The analysis of future climate projections from numerical climate model
simulations is of paramount importance to assess future climate change under
different forcing scenarios. Since this involves metrics of public interest
like the allowable fossil fuel emissions to meet particular warming targets,
for example the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
\autocite{UNFCCC2015}, research in this field of science is not only relevant
to climate scientists but to policymakers and the whole human society. In the
light of the large spread of climate sensitivity in the most recent generation
of climate models from \acs{CMIP}6, that has been demonstrated in
\cref{ch:04:papers_ecs_tcr_assessment}, a careful statistical evaluation and
refinement of the output of multi-model climate projections is as relevant as
ever. This thesis quantifies the associated uncertainties, presents the
evaluation of established methods to reduce these uncertainties in the new
climate model ensemble, and finally introduces a novel alternative technique
based on supervised \ac{ML}.

@@

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
\acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
on the upper and lower end than any model range from previous \ac{CMIP}
generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of
\ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus
$3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
\acs{CMIP}6 models range is considerably larger than in the assessed range of
$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
\unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with many models
showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model range of
\acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the \acs{CMIP}5
range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased climate
sensitivity in some \acs{CMIP}6 models is the addition of prognostic aerosol
schemes that include aerosol-cloud interactions, which might result in overly
large negative radiative forcing. In this case, a stronger response to \ac{GHG}
forcing is required to correctly reproduce the observed historical temperature
trend, resulting in a higher climate sensitivity in the corresponding models. A
further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in the
microphysical representation of mixed-phase clouds in some \acs{CMIP}6 models.
This change was implemented to improve the representation of the fractioning of
cloud ice and cloud liquid in those models, \eg{} by allowing for (more)
supercooled cloud liquid water. In addition, the \acs{CMIP}6 \ac{MMM} shows an
improved simulation of the shortwave \ac{CRE} when compared to satellite
observations. The change in the representation of mixed-phase clouds in some
\acs{CMIP}6 models reduces the strong negative shortwave cloud feedback over
the Southern Ocean that is present in predecessor versions of these models and
that results from a cloud phase change from ice clouds in the present-day to
liquid clouds in the future. In the affected \acs{CMIP}6 models, this cloud
phase change due to warming is reduced since these models simulate less cloud
ice over the present-day Southern Ocean than their predecessor versions with no
supercooled cloud liquid.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, already-published emergent constraints that have been derived
on models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5
are evaluated for their skill in the \acs{CMIP}6 ensemble. Emergent constraints
use a physically-based inter-model relationship between an observable quantity
of the Earth system and a target variable to reduce uncertainties in the target
variable with observations \autocite{Allen2002}. In total eleven emergent
constraints on \ac{ECS} are assessed, which are mostly related to cloud
feedbacks since these constitute the most important source of uncertainty for
\ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated emergent
constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5 ensemble,
out-of-sample tests on the emergent constraints can be performed by assessing
whether they still hold for the \acs{CMIP}6 models. In this study, which is
shown in \cref{ch:05:paper_ecs} and already published in
\textcite{Schlund2020a}, a substantial reduction of skill for the majority of
emergent constraints is found when applied to the \acs{CMIP}6 ensemble in
comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
decrease of the coefficient of determination $R^2$ of the emergent relationship
and a decrease of the statistical significance using the null hypothesis that
there is no correlation between the predictor and \ac{ECS}. Moreover, the
corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
emergent constraints. Thus, \cref{enum:01:question_2}
(\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{not
  very well} for the \acs{CMIP}6 ensemble. The increased best estimates and
spreads resulting from the emergent constraints in \acs{CMIP}6 are likely
related to the increased \ac{MMM} and multi-model spread of \ac{ECS} in
\acs{CMIP}6. A possible reason for the reduced skill of the emergent
constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \Ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, the global
mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario is constrained to $\pmrangeunit{171}{12}{\GtCyr{}}$ using
a published emergent constraint by \textcite{Wenzel2016}. In contrast to that,
the corresponding multi-model range of \ac{GPP} in the raw \acs{CMIP}5 ensemble
is $\rangeunit{156}{247}{\GtCyr{}}$. This first step corrects the \acs{CMIP}5
models' \ac{GPP} response to \ac{CO2}, which is primarily driven by the
\ac{CO2} fertilization. In the second step, a \ac{ML}-based climate model
weighting approach is used to further constrain the gridded \ac{GPP} based on
present-day predictors that are relevant for the simulation of \ac{GPP} in the
\acp{ESM}. The \ac{ML} approach is mathematically similar to the \ac{MDER}
approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}, but
additionally considers multi-dimensional (gridded) target variables and
non-linear relationships between the predictors and the target variable. A a
relationship between process-oriented predictors and future projections of
\ac{GPP} is established and then utilized to project today's observed
conditions into the future. The prediction phase of the new method can be
interpreted as an implicit performance weighting. However, due to the complex
structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
extract specific values for the individual weights. Two target variables are
considered: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave-one-model-out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
and \ac{LAI} are the most important features. A corresponding feature map shows
that \ac{T} dominates in the tropics and \ac{LAI} dominates in the northern
high latitudes, which might be attributed to an additional greening trend in
the high latitudes that enhances the \ac{GPP} increase in this regions compared
to regions closer to the equator through an extension of the growing season.
This study directly addresses \cref{enum:01:question_3}
(\enquote{\KeyScienceQuestionThree{}}), which can be confidently answered with
\enquote{yes} based on the results found.

@@

Climate sensitivity is a policy-relevant and easy-to-use metric to assess the
strength of climate change. The range of model results for this important
climate metric, however, remains large and has not been narrowed over the last
decades. In the latest generation of climate models contributing to
\acs{CMIP}6, the range has even increased \autocite{Meehl2020}.
\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
changes in the aerosol-cloud interaction and changes in the shortwave cloud
phase change feedback over the Southern Ocean. While these are likely
explanations, there might be additional relevant aspects that have not been
analyzed in detail yet. This could include additional feedback processes that
are not present in older \ac{CMIP} generations. Obvious candidates for these
feedback processes are cloud-related feedbacks, which pose a major source of
uncertainty in climate sensitivity in modern-day \acp{ESM}
\autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
can further help to gain a better understanding of the \acs{CMIP}6 models and
potentially to reduce associated uncertainties in the entire multi-model
ensemble. A potentially interesting process relevant in this context is the
influence of the midlatitude jet position on clouds and \acp{CRE}
\autocite{Grise2016}.

>>>>

The analysis of future climate projections from numerical climate model
simulations is of paramount importance to assess future climate change under
different forcing scenarios. Since this involves metrics of public interest
like the allowable fossil fuel emissions to meet particular warming targets,
for example the $1.5 \unit{\degreeCelsius}$ of the Paris Agreement
\autocite{UNFCCC2015}, research in this field of science is not only relevant
to climate scientists but to policymakers and human society as a whole. In the
light of the large spread of climate sensitivity in the most recent generation
of climate models from \acs{CMIP}6, that has been demonstrated in
\cref{ch:04:papers_ecs_tcr_assessment}, a careful statistical evaluation and
refinement of the output of multi-model climate projections is as relevant as
ever. This thesis quantifies uncertainties in climate model projections,
presents the evaluation of established methods to reduce these uncertainties in
a new climate model ensemble, and finally introduces a novel alternative
technique based on supervised \ac{ML}.

@@

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS}, a
\acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found, which is higher
on the upper and lower bound than any model range from previous \ac{CMIP}
generations. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of \ac{ECS}
is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus $3.23
\unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
\acs{CMIP}6 models range is considerably larger than in the assessed range of
$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
\unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with many models
showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model range of
\acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the \acs{CMIP}5
range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from \acs{AR}5 of
$\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased climate
sensitivity in some \acs{CMIP}6 models is the addition of prognostic aerosol
schemes that include aerosol-cloud interactions, which might result in overly
large negative radiative forcing. In this case, a stronger response to \ac{GHG}
forcing is required to correctly reproduce the observed historical temperature
trend, resulting in a higher climate sensitivity in the corresponding models. A
further reason for the high \ac{ECS} in \acs{CMIP}6 is a change in the
microphysical representation of mixed-phase clouds in some \acs{CMIP}6 models.
This change was implemented to improve the representation of the fractioning of
cloud ice and cloud liquid in those models, \eg{} by allowing for (more)
supercooled cloud liquid water. In addition, the \acs{CMIP}6 \ac{MMM} shows an
improved simulation of the shortwave \ac{CRE} when compared to satellite
observations. The change in the representation of mixed-phase clouds in some
\acs{CMIP}6 models reduces the strong negative shortwave cloud feedback over
the Southern Ocean that is present in predecessor versions of these models and
that results from a cloud phase change from ice clouds in the present-day to
liquid clouds in the future. In the affected \acs{CMIP}6 models, this cloud
phase change due to warming is reduced since these models simulate less cloud
ice over the present-day Southern Ocean than their predecessor versions with no
supercooled cloud liquid.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, already-published emergent constraints that have been derived
from models from the previous \ac{CMIP} generations \acs{CMIP}3 and \acs{CMIP}5
are evaluated for their skill in the \acs{CMIP}6 ensemble. Emergent constraints
use a physically-based inter-model relationship between an observable quantity
of the Earth system and a target variable to reduce uncertainties in the target
variable with observations \autocite{Allen2002}. In total eleven emergent
constraints on \ac{ECS} are assessed, which are mostly related to cloud
feedbacks since these constitute the most important source of uncertainty for
\ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated emergent
constraints have been derived from the \acs{CMIP}3 or \acs{CMIP}5 ensemble,
out-of-sample tests on the emergent constraints can be performed by assessing
whether they still hold for the \acs{CMIP}6 models. In this study, which is
shown in \cref{ch:05:paper_ecs} and already published in
\textcite{Schlund2020a}, a substantial reduction of skill for the majority of
emergent constraints is found when applied to the \acs{CMIP}6 ensemble in
comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
decrease of the coefficient of determination $R^2$ of the emergent relationship
and a decrease of the statistical significance using the null hypothesis that
there is no correlation between the predictor and \ac{ECS}. Moreover, the
corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
emergent constraints. Thus, \cref{enum:01:question_2}
(\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{not
  very well} for the \acs{CMIP}6 ensemble. The increased best estimates and
spreads resulting from the emergent constraints in \acs{CMIP}6 are likely
related to the increased \ac{MMM} and multi-model spread of \ac{ECS} in
\acs{CMIP}6. A possible reason for the reduced skill of the emergent
constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
for the latest generation of \acp{ESM} from \acs{CMIP}6 due to an increased
number of processes that are included in these models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. Therefore, this analysis does not focus on
reducing uncertainties in climate sensitivity itself but rather on a selected
process that contributes to it: \ac{GPP}. \Ac{GPP} is the largest flux of the
terrestrial carbon uptake and slows down global warming by removing \ac{CO2}
from the atmosphere. In the first step of the new two-step approach, a
published emergent constraint by \textcite{Wenzel2016} is used to constrain the
global mean \ac{GPP} at the end of the \nth{21} century in the \acs{CMIP}5
\acs{RCP}8.5 scenario to $\pmrangeunit{171}{12}{\GtCyr{}}$, compared to
unconstrained \acs{CMIP}5 model range of $\rangeunit{156}{247}{\GtCyr{}}$. This
first step corrects the \acs{CMIP}5 models' \ac{GPP} response to \ac{CO2},
which is primarily driven by the \ac{CO2} fertilization. In the second step, a
\ac{ML}-based climate model weighting approach is used to further constrain the
gridded \ac{GPP} based on present-day predictors that are relevant for the
simulation of \ac{GPP} in the \acp{ESM}. The \ac{ML} approach is mathematically
similar to the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020,
  Wenzel2016a}, but additionally considers multi-dimensional (gridded) target
variables and non-linear relationships between the predictors and the target
variables. A relationship between process-oriented predictors and future
projections of \ac{GPP} is established and then utilized to project today's
observed conditions into the future. The prediction phase of the new method can
be interpreted as an implicit performance weighting. However, due to the
complex structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible
to extract specific values for the individual weights. Two target variables are
considered: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. The new approach is validated by
comparing it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave-one-model-out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) is found when using the \ac{ML} approach. Moreover,
the evaluation of the global and local feature importance allows further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by the
new approach. For the second variable (fractional change in \ac{GPP}), \ac{T}
and \ac{LAI} are the most important features. A corresponding feature map shows
that \ac{T} dominates in the tropics and \ac{LAI} dominates in the northern
high latitudes, which might be attributed to an additional greening trend in
the high latitudes that enhances the \ac{GPP} increase in this regions compared
to regions closer to the equator through an extension of the growing season.
This study directly addresses \cref{enum:01:question_3}
(\enquote{\KeyScienceQuestionThree{}}), which can be confidently answered with
\enquote{yes} based on the results found.

@@

Climate sensitivity is a policy-relevant and easy-to-use metric to assess the
strength of climate change. The range of model results for this important
climate metric, however, remains large and could not be narrowed down over the
last decades. In the latest generation of climate models contributing to
\acs{CMIP}6, the range has even increased \autocite{Meehl2020}.
\Cref{ch:04:papers_ecs_tcr_assessment} presents two possible reasons for this:
changes in the aerosol-cloud interaction and changes in the shortwave cloud
phase change feedback over the Southern Ocean. While these are likely
explanations, there might be additional relevant aspects that have not been
analyzed in detail yet. This could include additional feedback processes that
are not present in older \ac{CMIP} model generations. Obvious candidates for
these feedback processes are cloud-related feedbacks, which pose a major source
of uncertainty in climate sensitivity in modern-day \acp{ESM}
\autocite{Boucher2013}. Identifying and quantifying such feedback mechanisms
can further help to gain a better understanding of the \acs{CMIP}6 models and
potentially to reduce associated uncertainties in the entire multi-model
ensemble. A potentially interesting process relevant in this context is the
influence of the midlatitude jet position on clouds and \acp{CRE}
\autocite{Grise2016}.
